---
title: "Weibull"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
editor_options: 
  chunk_output_type: inline
---

## (Old, from 10/10/2019) Conditional Survival 

This Weibull parameterization has pdf $$f(t| \lambda, \nu) = \nu \lambda t^{\nu - 1} exp(-\lambda t^{\nu}),$$
and survival function
$$S(t) =  exp(-\lambda t^\nu).$$

Use the properties 
$$\Lambda(t) = \int_{0}^{t} \lambda(t)$$
and 
$$S(t) = exp(-\Lambda(t)),$$
where $\lambda(t)$ is the hazard function at time $t$, $\Lambda(t)$ is the cumulative hazard function, and $S(t)$ is the survival function. Then 

$$S(t|t_0) = exp\{-\int_{t_0}^{t + t_0} \lambda(x)dx\} = exp\{-\int_{0}^{t + t_0} \lambda(x)dx -\int_{0}^{t_0} \lambda(x)dx \}  = \frac{S(t + t_0)}{S(t_0)}$$ 

Plug in the weibull survival functions: 
$$ S(t|t_0) = \frac{S(t + t_0)}{S(t_0)} = \frac{exp(\lambda (t + t_0)^\nu)}{exp(\lambda t_0^\nu)}  =  exp[ \lambda \{ (t + t_0)^\nu - t_0^\nu\}]$$

Then the conditional CDF is $$F(t|t_0) = 1 - S(t|t_0) = 1 - exp[ \lambda \{ (t + t_0)^\nu - t_0^\nu\}]$$

Invert to get the quantile function. After a lot of algebra this expression $u = 1 - exp[ \lambda^\nu \{ (t + t_0)^\nu - t_0^\nu\}]$ becomes 
$$F^{-1}(u) = \{{t_0}^\nu - \lambda^{-1}log(1-u) \}^{1/\nu} - t_0$$

R function:

```{r}
f_inv_cond_weib = function(u, nu, lambda, t0){
  
  t = (t0^nu - log(1-u) * lambda^(-1))^(1/nu) - t0
  
  return(t)
  
}
```


## Odds Adjusted Conditional Survival

Under a  proportional hazards assumption we have

$$S_1(t) = S_0(t)^\beta, $$

where $S_0(\cdot)$ is the survival function in the reference group, $S_1(\cdot)$ is the survival function in the treatment group, and $\beta$ is the hazard ratio.

Plug in $S_1(t)$ into the expression for the conditional survival function above to get

$$S_1(t | t_0) = \bigg[ \frac{S_0(t + t_0)}{S_0(t_0)} \bigg]^\beta = exp[ \lambda \{ (t + t_0)^\nu - t_0^\nu\}]^\beta = exp[ \beta \lambda \{ (t + t_0)^\nu - t_0^\nu\}]$$

Then the conditional survival function is 

$$F(t|t_0) = 1 -  exp[ \beta \lambda \{ (t + t_0)^\nu - t_0^\nu\}].$$

Inverting this gets

$$Q(u | t_0) = \{{t_0}^\nu - (\lambda \beta)^{-1} log(1-u) \}^{1/\nu} - t_0.$$


R function:

```{r}
# The hazard ratio is set to 1 by default so the function can be used for 
# non hazard adjusted inverse CDF's
inverse_cond_cdf = function(u, nu, lambda, beta_HR = 1, t0){
  
  t = (t0^nu - log(1 - u) * (beta_HR * lambda)^(-1) )^(1/nu) - t0
  
  return(t)
  
}


curve(inverse_cond_cdf(x, nu = 1, lambda = 0.5, 
                       beta_HR = 2, t0 = 1),
      from = 0, 
      to = 1,
      xlab = "Quantile",
      y = "",
      main = "Inverse CDF for Weibull Model Adjusted for Hazard Ratio",
      lwd = 3
)
```
